Variana Reloaded

### 2022-11-03 cv/cv.tex
Moving to canonical coordinates explained in variana notes

\begin{flushright} 
{\huge {\textbf{Alexis Roche}}}\\
\begin{small}
+351 910 327 756 $\mid$ \url{alexis.roche@gmail.com}\\
\url{https://www.linkedin.com/in/alexis-roche-3815592}\\
%French National $\mid$ DOB 12/8/1972 $\mid$ Married, 2 Children\\
French National $\mid$ Married, 2 Children\\
\end{small}
\end{flushright}

@@

\noindent Hands-on scientific team leader, graduated from \'Ecole Centrale Paris, PhD in engineering science, with a specialist expertise in computer vision and machine learning. Has operated in academic research for 15 years before moving to R\&D in industry (healthcare, electro-domestic, entertainment). Strong experience in the development of both open source and commercial software in Python, C, C++. {\em Keywords:} applied mathematics, statistics, artificial intelligence, machine learning, computer vision, medical image analysis, clinical studies, computer graphics.

@@

\noindent \textbf{STATISTICAL MODELING:} Strong experience developing model-based algorithms for statistical inference, with applications in machine learning, computer vision, computer graphics, \ldots

\ \\
\noindent \textbf{TECHNICAL LEADERSHIP:} Proven ability to coordinate a team and manage objectives in multidisciplinary environments.

\ \\
\noindent \textbf{SCIENTIFIC COMPUTING:} 15 years experience in Python programming within the NumPy/SciPy ecosystem and interfacing C~language with Python.

\ \\
\noindent \textbf{SCIENTIFIC WRITING:} Main author of 25 peer-reviewed scientific articles published in international journals, proceedings and books, co-author of more than 100 scholarly publications and 5 published international patents.

@@

\bighead{HEAD OF COMPUTER VISION \& ARTIFICIAL INTELLIGENCE}{Didimo, Porto, Portugal}{2019--present}
\nopagebreak
\begin{itemize}
\item Leading development of algorithms for automated generation of high-fidelity digital humans from photos. {\em Keywords:} facial reconstruction, ``deep face'', animation retargetting, image enhancement.
\end{itemize}
\ \\

@@

\ \\
\bighead{PERMANENT RESEARCHER -- NEUROIMAGING INSTITUTE}{French Atomic Commission (CEA), Paris, France}{2002--2011}
\nopagebreak
\begin{itemize}
\item Developed advanced algorithms for image processing and statistical analysis of brain imaging data. {\em Keywords:} functional, anatomical and diffusion-weighted MRI, image-based population analysis, spatio-temporal image registration, real-time imaging.
\item Funded project management: principal investigator ({\em Karametria} -- statistical analysis of brain structures, French National Research Agency project, budget: 620 K\euro, 2009--2011); team leader ({\em NIBB} -- studying language in infants via functional neuroimaging, French National Research Agency project, 2006--2009).
\item Active contributor to the NiPy software library (Neuroimaging in Python, \url{www.nipy.org}) from 2006.
\item Academic Guest at Computer Vision Laboratory, Swiss Federal Institute of Technology Zurich (ETHZ), Switzerland, 2009--2011.
\end{itemize}

@@

\begin{tabular}{p{0.2\linewidth}p{0.8\linewidth}}
  \textbf{Languages:} & Native French, fluent English, basic Italian \& Portuguese.\\
  \hline
  \textbf{IT skills:} &
  \tabitem Programming languages: Python, C, C++, Matlab, R.\\
  & \tabitem Long-term experience with scientific Python pacakges (numpy, scipy, pylab).\\
  & \tabitem CV/AI Python packages: skimage, PIL, opencv, sklearn, pytorch, tensorflow, keras.\\
  & \tabitem  C/Python integration via Cython.\\
  & \tabitem Version control software: git, svn, perforce.\\
  & \tabitem Agile workflow (mainly Scrum) using JIRA \& Confluence.\\
  & \tabitem Reporting: \LaTeX, MS~Office, OpenOffice, LibreOffice, Google~Docs.\\
  \hline
  \textbf{Publications:} &
  \tabitem ORCID: \url{http://orcid.org/0000-0002-4821-6893}. \\
  & \tabitem 115 scholarly publications.\\
  & \tabitem Google scholar statistics: 6862 citations, h-index: 33, i10-index: 68 (March 2022).\\
  & \tabitem 5 published international patents. 2 ongoing applications.\\
  \hline
  \textbf{Academic work:} & 
  \tabitem Post-graduate teaching in computer vision and medical image analysis: about 100~hours (Ecole Centrale Paris, EPFL Lausanne, University of Nice, University of Lausanne, INSERM/CNRS continuing education, \ldots).\\
  & \tabitem Long-time scientific journal reviewing experience (IEEE Trans.~Medical Imaging, Medical Image Analysis, NeuroImage, IEEE Trans.~PAMI, IEEE Trans.~Signal Processing, Frontiers in Neuroscience, \ldots).\\
  & \tabitem Supervision: two PhD students and four MSc students as main supervisor, three PhD students and two MSc students as co-supervisor.\\
  \hline
  \textbf{References:} & Available on request.
\end{tabular}

>>>>

\begin{flushright} 
{\huge {\textbf{Alexis Roche}}}\\
\begin{small}
+351 910 327 756 $\mid$ \url{alexis.roche@gmail.com}\\
\url{https://www.linkedin.com/in/alexis-roche-3815592}\\
%French National $\mid$ DOB 12/8/1972 $\mid$ Married, 2 Children\\
French National $\mid$ Married, 2 Children\\
Based in Porto, Portugal.
\end{small}
\end{flushright}

@@

\noindent Hands-on technical team leader, engineer at heart, PhD, with a specialist expertise in artificial intelligence and data analysis. Has operated in public research and private R\&D in various fields (healthcare, electro-domestic, entertainment). Strong experience in the development of both open source and commercial software in Python, C, C++. {\em Keywords:} applied mathematics, statistics, machine learning, computer vision, computer graphics, medical image analysis, clinical studies.

@@

\noindent

\noindent \textbf{TECHNICAL LEADERSHIP:} Proven ability to coordinate a team and manage objectives in multidisciplinary environments.

\ \\
\noindent \textbf{SCIENTIFIC PROGRAMMING:} 15 years experience programming in Python and~C language in the SciPy ecosystem.

\ \\
\textbf{STATISTICS:} Strong experience developing inferential statistical models for machine learning, computer vision, computer graphics, and performing data analysis for cognitive research, clinical studies, software validation, \ldots

\ \\
\noindent \textbf{SCIENTIFIC REPORTING:} Main author of 25 peer-reviewed scientific articles published in international journals, proceedings and books, co-author of more than 100 scholarly publications and 5 published international patents.

@@

\bighead{HEAD OF DIDIMO AUTOMATION TEAM}{Didimo, Porto, Portugal}{2019--present}
\nopagebreak
\begin{itemize}
\item Leading development of generation pipeline of high-fidelity avatars from photos. {\em Keywords:} facial reconstruction, ``deep face'', animation retargetting, image feature extraction, image restoration.
\end{itemize}
\ \\

@@

\ \\
\bighead{PERMANENT RESEARCHER -- NEUROSPIN (NEUROIMAGING INSTITUTE)}{French Atomic Commission (CEA), Paris, France}{2002--2011}
\nopagebreak
\begin{itemize}
\item Developed advanced algorithms for image processing and statistical analysis of brain imaging data. {\em Keywords:} functional, anatomical and diffusion-weighted MRI, image-based population analysis, spatio-temporal image registration, real-time imaging.
\item Funded project management: principal investigator ({\em Karametria} -- statistical analysis of brain structures, French National Research Agency project, budget: 620 K\euro, 2009--2011); team leader ({\em NIBB} -- studying language in infants via functional neuroimaging, French National Research Agency project, 2006--2009).
\item Active contributor to the NiPy software library (Neuroimaging in Python, \url{www.nipy.org}) from 2006.
\item Academic Guest at Computer Vision Laboratory, Swiss Federal Institute of Technology Zurich (ETHZ), Switzerland, 2009--2011.
\end{itemize}

@@

\begin{tabular}{p{0.2\linewidth}p{0.8\linewidth}}
  \textbf{Languages:} & Native French, fluent English, basic Portuguese \& Italian.\\
  %\hline
  \ \\
  \textbf{IT skills:} &
  \tabitem Programming languages: Python, C, C++, Matlab, R.\\
  & \tabitem Long-term experience with scientific Python packages (numpy, scipy, pylab).\\
  & \tabitem CV/AI Python packages: skimage, PIL, opencv, sklearn, pytorch, tensorflow, keras.\\
  & \tabitem  C/Python integration via Cython.\\
  & \tabitem Version control software: git, svn, perforce.\\
  & \tabitem Agile workflow (mainly Scrum) using JIRA \& Confluence.\\
  & \tabitem Reporting: \LaTeX, MS~Office, OpenOffice, LibreOffice, Google~Docs.\\
  %\hline
  \ \\
  \textbf{Publications:} &
  \tabitem ORCID: \url{http://orcid.org/0000-0002-4821-6893}. \\
  & \tabitem 116 scholarly publications.\\
  & \tabitem Google scholar statistics: 7308 citations, h-index: 35, i10-index: 72 (October 2022).\\
  & \tabitem 5 published international patents. 2 ongoing applications.\\
  %\hline
  \ \\
  \textbf{Academic work:} & 
  \tabitem Post-graduate teaching in computer vision and medical image analysis: about 100~hours (France: Ecole Centrale Paris, University of Nice, INSERM/CNRS continuing education; Switzerland: EPFL Lausanne, University of Lausanne).\\
  & \tabitem Long-time scientific journal reviewing experience (IEEE Trans.~Medical Imaging, Medical Image Analysis, NeuroImage, IEEE Trans.~PAMI, IEEE Trans.~Signal Processing, Frontiers in Neuroscience, \ldots).\\
  & \tabitem Supervision of two PhD students and four MSc students as main supervisor, three PhD students and two MSc students as co-supervisor.\\
  %\hline
  \ \\
  \textbf{References:} & Available on request.
\end{tabular}

### 2022-11-03 papers/variana/notes.tex
Updates summer 2022
Moving to canonical coordinates explained in variana notes

@@

So, let's say that we have a current fit $g_n$ (which may be defined, for instance, by its coordinates in the canonical, fixed basis) and we want to perform one Newton iteration to improve it. Given an orthonormal basis $\phi_n$ associated with $g_n$ (in the sense of the above-defined dot product), we may look for the update under the form $g_\theta = g_n \exp({\theta^\top \phi_n})$. The Newton iteration from $\theta=0$ then takes the form:
$$
\hat{\theta} = -\lambda \nabla D(0) = \lambda \int \pi(f - g_n)\phi_n,
$$
for some damping factor $\lambda > 0$ (the genuine Newton iteration corresponds to $\lambda=1$), yielding the update:
$$
g_{n+1} = g_n \exp (\hat{\theta}^\top \phi_n).
$$

We will further need to represent $g_{n+1}$ in the same fixed coordinate system as $g_n$, which will require a linear coordinate change from $\phi_n$ to the fixed basis.

@@

Given some ``moving'' coordinates in the orthonormal basis, we aim to compute the corresponding coordinates in the canonical basis $b_0=1$, $b_{11}=x_1$, $b_{12}=x_1^2$, $b_{21}=x_2$, $b_{22}x_2^2$, etc. This is a linear mapping. We easily get by identification, owing to the separability of Gaussian factors:
$$
\phi_0 =
\frac{1}{\sqrt{Z}} b_0,
\qquad
\phi_{1i} =
-\frac{m_i}{\sqrt{Z}s_i} b_0
+ \frac{1}{\sqrt{Z}s_i} b_{1i},
\qquad
\phi_{2i} =
\frac{1}{\sqrt{2Z}} \left(\frac{m_i^2}{s_i^2} - 1\right) b_0
- \frac{\sqrt{2}m_i}{\sqrt{Z}s_i^2} b_{1i}
+ \frac{1}{\sqrt{2Z}s^2} b_{2i}.
$$

This shows that it the transition matrix from moving to fixed coordinates is block-diagonal, up to a first row corresponding to the constant terms. In the 1D case, it is upper triangular. In general, it never has more than three nonzero elements per column.

@@

>>>>

Note that this is only tight if $\max(f,g)$ is constant,

@@

So, let's say that we have a current fit $g_n$ (which may be defined, for instance, by its coordinates in the canonical, fixed basis) and we want to perform one Newton iteration to improve it. Given an orthonormal basis $\phi_n$ associated with $g_n$ (in the sense of the above-defined dot product), we may look for the update under the form $g_\nu = g_n \exp({\nu^\top \phi_n})$. The Newton iteration from $\nu=0$ then takes the form:
$$
\hat{\nu} = -\lambda \nabla D(0) = \lambda \int \pi(f - g_n)\phi_n,
$$
for some damping factor $\lambda > 0$ (the genuine Newton iteration corresponds to $\lambda=1$), yielding the update:
$$
g_{n+1} = g_n \exp (\hat{\nu}^\top \phi_n).
$$

We will further need to represent $g_{n+1}$ in the same fixed coordinate system as $g_n$, i.e. using a ``$\theta$ parameter'', which will require a linear coordinate change from $\phi_n$ to the fixed basis.

@@

Given some ``moving'' coordinates in the orthonormal basis, we aim to compute the corresponding coordinates in the canonical basis $b_0=1$, $b_{11}=x_1$, $b_{12}=x_1^2$, $b_{21}=x_2$, $b_{22}x_2^2$, etc. This is a linear mapping. We easily get by identification, owing to the separability of Gaussian factors:
$$
\phi_0 =
\frac{1}{\sqrt{Z}} b_0,
\qquad
\phi_{1i} =
-\frac{m_i}{\sqrt{Z}s_i} b_0
+ \frac{1}{\sqrt{Z}s_i} b_{1i},
\qquad
\phi_{2i} =
\frac{1}{\sqrt{2Z}} \left(\frac{m_i^2}{s_i^2} - 1\right) b_0
- \frac{\sqrt{2}m_i}{\sqrt{Z}s_i^2} b_{1i}
+ \frac{1}{\sqrt{2Z}s_i^2} b_{2i}.
$$

This shows that the transition matrix from moving to fixed coordinates is block-diagonal, up to a first row corresponding to the constant terms. In the 1D case, it is upper triangular. In general, it never has more than three nonzero elements per column.

$$
T =
\frac{1}{\sqrt{Z}}
\left[
\begin{array}{ccccccc}
1 & -\frac{m_1}{s_1} & -\frac{m_2}{s_2} & \ldots &
  \frac{1}{\sqrt{2}} \left(\frac{m_1^2}{s_1^2} - 1\right) &
  \frac{1}{\sqrt{2}} \left(\frac{m_2^2}{s_2^2} - 1\right) &
  \ldots \\
0 & \frac{1}{s_1} & 0 & \ldots &
-\frac{\sqrt{2}m_1}{s_1^2} & 0 & \ldots \\
  0 & 0 & \frac{1}{s_2} & 0 &
  0 & -\frac{\sqrt{2}m_2}{s_2^2} & \ldots \\
\vdots & \vdots & \vdots & \vdots &
\vdots & \vdots & \vdots\\
0 & 0 & 0 & 0 &
  \frac{1}{\sqrt{2}s_1^2} & 0 & \ldots\\
0 & 0 & 0 & 0 &
  0 & \frac{1}{\sqrt{2}s_2^2} & \ldots\\
\end{array}
\right].
$$

@@

In VS, we approximate the KL divergence using some integration rule with weights~$w_i$:
$$
\hat{D}(\nu)
\equiv \sum_i w_i \left[-f_i
  \log g_\nu(\x_i) + g_\nu(\x_i) \right],
$$
yielding the gradient (with simplified notation):
$$
\nabla\hat{D} =
\sum_i w_i (g_i - f_i) \phi(\x_i),
$$
where $\phi$ is the above defined orthonormal basis, which depends on the current fit~$g$.

Hence, the VS approximation to the Newton iteration on the KL~divergence is:
$$
\hat{\nu} = -\lambda \sum_i w_i (g_i - f_i) \phi(\x_i).
$$

This requires evaluating both the current fit and the associated orthonormal basis functions at each sample point. Note that, in the canonical basis, we don't have to re-evaluate the basis functions on each iteration, but we have to re-evaluate the KL Hessian if we are to implement the genuine Newton method.

Once the moving coordinates are updated through $\hat{\nu}$, we convert them to canonical coordinates by left multiplying $\hat{\nu}$ by the above transition matrix.

\subsection{Utter bulshit}
